---
#title: "Reproducible Evidence Synthesis"
#subtitle: "How to lit review"  
#author: "Yacob Haddou"
#date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    seal: false
    nature:
      titleSlideClass: ["left"]
      slideNumberFormat: "%current%/%total%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.retina=2,
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE,
  eval=FALSE,
  tidy=TRUE
)

library(xaringan); library(fontawesome); library(tidyverse)
```

```{r xaringan-tile-view, echo=FALSE}
xaringanExtra::use_tile_view()
```

```{r xaringan-scribble, echo=FALSE}
xaringanExtra::use_scribble()
```

```{r xaringanExtra, echo = FALSE}
xaringanExtra::use_progress_bar(color = "#ffffff", location = "bottom")
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#264653",
  secondary_color = "#ce4257",
  header_font_google = google_font("Quicksand Medium"),
  text_font_google = google_font("Quicksand Medium"),
  code_font_google = google_font("Fira Pro"),
  text_font_size = "150%"
  )
```

.right[
# Reproducible Evidence Synthesis
## How to lit review  
]

----

.left[

>##### [`r fa_i("envelope")` yacob.haddou@glasgow.ac.uk](mailto:yacob.haddou@glasgow.ac.uk)<br>

>##### [`r fa_i("user-ninja")` https://valiriel.github.io](mailto:yacob.haddou@glasgow.ac.uk)<br>

>##### [`r fa_i("github")` `r fa_i("researchgate")` .../valiriel](https://github.com/valiriel)

>##### [`r fa_i("orcid")` https://orcid.org/0000-0003-0557-0339](https://orcid.org/0000-0003-0557-0339)

]

---

## Disclaimer & context
--------------

- I have not invented this, just reshaped it to fit my needs

- This is part of my PhD initial review, <b> first time </b> I do an "official" review

##### _ Zhi Foo et al. _  (2021) A practical guide to question formation, systematic searching and study screening for literature reviews in ecology and evolution. Methods Ecol. Evol. DOI: 10.1111/2041-210X.13654

##### _ Grames et al. _  (2019) An automated approach to identifying search terms for systematic reviews using keyword co‐occurrence networks. Methods Ecol. Evol. DOI: 10.1111/2041-210X.13268

---

## Start somewhere...
------------------------

- Come up with your unambiguous research question, clearly defined.

- Get an idea of the _ high citation _ articles on your topic.

- <b> Seek out any review already done on your field ! </b> 

- Avoid overlaps with what has already been synthesized before by clarifying the timeframe, aim and search string of past reviews. 

- Now come up with a naive search question.

---

## Naive search question
-------------------------

1. Come up with a series of keywords that represent your question.

  - PICO method (Population, Intervention, Control, Outcome)
  - Think of it as Study system, Predictor variables and Response
  - Consider lateral/synonyms and vertical/specific/overarching terms
  - Check with peers that it all makes sense

2. Use boolean operator <b>(AND, OR, ...)</b> to put the terms together 

3. Use truncation/stemming <b>(*)</b> to capture word with common root and plurals, be cautious though

---

## Example naive search question
--------------------------------

After having a look at past publications/reviews and what said above I came up with this series of term.

Not interested in a specific response or explanatory so kept it wide for anything including these terms.

.inverse[ 

##### ("extinction debt/" OR "species credit/" OR "colonization credit/" OR "colonisation credit/" OR "immigration credit/" OR "ecosystem function debt/" OR "ecosystem service debt/" OR "ecosystem function credit/" OR "ecosystem service credit/" OR "ecological lags" OR "biodiversity lag/" OR "species lags" OR "landscape legacy") 

]

substitute / with * (markdown does not interpret otherwise)

---

## Time to do a search
----------------------

- Pick some literature databases and query them.
  - Web of Science, BASE, SCOPUS, ScienceDirect etc...

- <b>Do</b> have a look at how many entries are returned, in the many 1000s you are probably asking something too broad and no one has time to go through all of that, honestly. 

- Export the entries as .ris files

---

## Install some packages
------------------------

### https://rmetaverse.github.io/

```{r}
remotes::install_github("rmetaverse/metaverse")
remotes::install_github("elizagrames/litsearchr",
                        ref="main")

```

Packages aimed at evidence syntheses and meta-analyses

---

## Import naive search results
--------------------------    

```{r}
naive_import <- import_results(directory = "lit_data/00_naive_lit_data/naive_search_data", verbose = TRUE)
print(paste0("imported n. ", nrow(naive_import)))

```
910 records from my naive search  

---

## De-duplicate
-------------------------- 

Based on exact match and string similarity of title

```{r}
naive_deduplicated <- remove_duplicates(naive_import, 
                                        field = "title", method = "exact")
print(paste0("deduplicated_exact n. ", nrow(naive_deduplicated)))

naive_deduplicated <- remove_duplicates(naive_deduplicated, 
                                        field = "title", method = "string_osa")
print(paste0("deduplicated_string n. ", nrow(naive_deduplicated)))

```

810 and then 799 remaining, 111 records gone

---

## Extract keywords 
-------------------------- 

```{r}
rake_keywords <- 
  extract_terms(text = paste(naive_deduplicated$title, naive_deduplicated$abstract, 
                             naive_deduplicated$keywords),
                method = "fakerake", 
                min_freq = 10, # at least 10 times 
                ngrams = TRUE, min_n = 2, # minimum 2 words keywords
                language = "English", 
                stopwords = get_stopwords("English")) %>% as_tibble()

# remove terms that are too broad, save to csv (ver1) 
# and then save for re-import (ver2)
write.csv(rake_keywords, file = "lit_data/00_naive_lit_data/naive_search_keywords_ver1.csv")
```

---

### Remove broad keywords and co-occurence network 
--------------------------

```{r}
#' obsolete or too broad terms manually removed, such as keywords related to specific 
#' habitat or taxonomic groups, broad words ("biodiversity", population size) and unrelated connecting words
rake_keywords <- read.csv("lit_data/00_naive_lit_data/naive_search_keywords_ver2.csv")[,-1]

#' create document-feature matrix
naive_dfm <- create_dfm(elements = paste(naive_deduplicated$title, 
                                         naive_deduplicated$abstract),
                        features = rake_keywords)

#' create a keyword co-occurrence network
naive_graph <- create_network(search_dfm = naive_dfm, min_studies = 5, min_occ = 5)

#' evaluate keyword importance and clean
nodes_rank <- strength(naive_graph)

terms_rank <- tibble(term=names(nodes_rank), strength=nodes_rank, row.names=NULL) %>%
                  mutate(rank=rank(strength, ties.method="min")) %>% arrange(strength)

```

---

## Enhance,enhance 
--------------------------

- Note that the final keyword order might not always be so useful, words such us biodiversity could be high ranking but not necessarily good.

#### Manually inspect the list and see if you find any keywords that could fit in your original naive string, if yes include them and remember to truncate.

---

## Final search string
--------------------------

.center[.inverse[#### ("extinction debt/" OR "species credit/" OR "colonization credit/" OR "colonisation credit/" OR "immigration credit/" OR "ecosystem function debt/" OR "ecosystem service debt/" OR "ecosystem function credit/" OR "ecosystem service credit/" OR "ecological lags" OR "biodiversity lag/" OR "species lag/" OR "landscape legac/")
.center[#+]
###"community lag/" OR "emigration debt/" OR "delayed species extinction/" OR "species debt/" OR "delayed species colonisation/" OR "delayed species colonization/" 
]]

---

## Loop one more time through the above steps
--------------------------

1. Search the lit databases with your final string
  - 966 entries retrieved

1. De-duplicate
  - 830 articles remaining

---

## Manual screening  (part 1)
--------------------------

```{r}
# save a copy of raw data, 830 articles
write.csv(deduplicated, file="lit_data/02_title_abstract_screening/00_lit_not_screened.csv")

# screen titles 
revtools::screen_titles(deduplicated) 
# manually move saved file from main dir to lit_data/02_title_abstract_screening as 01_lit_title_screened.csv
data <- read.csv("lit_data/02_title_abstract_screening/01_lit_title_screened.csv") 

# retain selection from above, remove excluded
data <- data %>% filter(screened_titles != "excluded") 
table(data$screened_titles); table(data$source_type); table(data$year)

# screen abstract
screen_abstracts(data %>% as.bibliography()) # manually open file above
# manually move saved file from main dir to lit_data/02_title_abstract_screening as 02_lit_abstract_screened.csv
```

---

## Manual screening (part 2)
--------------------------

```{r}
# import again
data2 <- read.csv("lit_data/02_title_abstract_screening/02_lit_abstract_screened.csv")

# retain selection from above, remove excluded
data2 <- data2 %>% filter(screened_abstracts != "excluded") 
table(data2$screened_abstracts); table(data2$source_type); table(data2$year)

# retain peer-reviewed articles
data2 <- data2 %>% filter(source_type == "JOUR")

# see trend through time
plot(names(table(data2$year)), as.numeric(table(data2$year)), type="l")

#' Go through each article and import into reference manager for reading
write.csv(data2, file="lit_data/02_title_abstract_screening/03_final_lit_screened.csv")
```

---

## Manual screening (part 3)
--------------------------

Some reference numbers:
- 830 to begin with
- 289 left after screening titles
- 208 left after screening abstracts
- 196 left after keeping only peer-reviewed material
- 190 after sourcing and final clean-up

---

## Make it reproducible and more
--------------------------

.middle[
.center[
####Show R code, GitHub, Zotero
]
]

---

## Reading time
--------------------------

Constructed a series of variables and criteria to quantitatively assess each article and then report

.center[
### Show csv
]

---

.left[

##### [`r fa_i("paper-plane")` ___ yacob.haddou@glasgow.ac.uk](mailto:yacob.haddou@glasgow.ac.uk)<br>
##### [`r fa_i("github")` ___ https://github.com/valiriel](https://github.com/valiriel)<br/>
##### [`r fa_i("orcid")` ___ https://orcid.org/0000-0003-0557-0339](https://orcid.org/0000-0003-0557-0339)<br/>
##### [`r fa_i("researchgate")` ___ https://www.researchgate.net/profile/Yacob-Haddou](https://www.researchgate.net/profile/Yacob-Haddou)<br/>

### slides powered by xaringan https://github.com/yihui/xaringan

]

